\Large \bf 3-Step, 1-Factor Authentication With Custom-Fit, In-Ear EEG

### 2017-02-07 paper.tex
References

\subsection*{Abstract}
In this paper, we present a system that provides 3-factors of authentication (knowledge, posession and inherence) in a single step, using brain-based authentication via a custom-fit, in-ear EEG. Across all subjects, we achieve a mean 0\% false acceptance and 0.12X\% false rejection rate with data from only one earpiece. In a preliminary test of an ``imposter'' spoofing attack, we find a 0\% false acceptance rate. Our results indicate that in-ear EEG could provide a discrete, convenient, multifactor authentication.

@@

In previous work, “one-step two-factor authentication” has been proposed as a
new approach to authentication that can provide the security benefits of two-
factor authentication without incurring the hassle costs of two-step verification.
By employing consumer-grade EEG (electroencephalogram) sensing
technologies, it was demonstrated in a 2013 passthoughts study that a user can
submit both a knowledge factor (i.e., secret thought) and an inherence factor
(i.e., brainwave signal unique to the individual) in a single step by performing a
single mental task. Additionally, the robustness of this method against
impersonation attacks was demonstrated, including conditions where the attacker
may have learned the target’s secret thought and/or secret task.

@@

\section{Related work}
\label{sec:org3732d7c}

The use of EEG as a biometric signal for user authentication has a short history.
In 2005, Thorpe et al. motivate and outline the design of a passthoughts system,
where, rather than typing a password, users authenticate 
by thinking of a passthought. Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp. In 2013, one group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor. In particular, the
lack of signal diversity from multiple EEG channels can be overcome by allowing
the users to choose their own personalized passthoughts (e.g., sing their favorite
song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\> \$10,000 USD)
and low usability (gel-based electrodes; aesthetic challenges of an EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task).

@@

Research in in-ear EEG is only several years old. Nonetheless, the concept has
attracted a lot of attention because of the discreetness factor of in-ear EEG over
traditional scalp-based EEG. A research team at the Imperial College London
and Aarhus University published a landmark paper in 2011 that introduced the
concept of in-ear EEG, demonstrating for the first time the feasibility of recording
brainwave signals from within the ear canal. Follow-up work from the same
group demonstrated its ability to produce signal-to-noise ratios comparable to
those from conventional EEG electrode placements, robustness to common
sources of artifacts, and use in a brain-computer interface (BCI) system based on
auditory evoked potentials and visual evoked potentials.

\underline{Citation to our past work} 
was the first to merge these two streams of work, using in-ear EEG signals for
user authentication with a consumer-grade device. United Sciences is currently
developing a consumer hearable called The Aware that will measure EEG from the ear.
Behavioral authentication methods such as keystroke dynamics and speaker
authentication can be categorized as one-step two-factor authentication
schemes. In both cases, the knowledge factor (password or passphrase) and
inherence factor (typing rhythm or speaker’s voice) are employed. In contrast, the
Nymi band supports one-step two-factor authentication via the inherence
factor (cardiac rhythm that is supposed to be unique to each individual) and the
possession factor (the wearing of the band on the wrist). However, as far as we
know, no one has proposed or demonstrated a one-step three-factor
authentication scheme.
\section{Methods}
\label{sec:org63d105c}
\subsection{{\bfseries\sffamily TODO} Manufacturing, materials}
\label{sec:orgb483219}
\subsection{Subjects}
\label{sec:orgaf30da9}
\subsection{{\bfseries\sffamily TODO} Tasks}
\label{sec:org86eea32}

@@

@@

We analyzed the EEG signals collected during the tasks using a support vector classifier (SVC). Since past work has shown that classification tasks in EEG-based BCI are linear \underline{cite}, we used XGBoost, a popular tool for generating ensemble classifiers for logistic linear classification \underline{cite}.

@@

Counter to our expectations, we found that referencing on the same side as the electrodes improves classification accuracy, compared to referencing on the other side of the head, as is commonly done in EEG. \underline{Theory as to why?} Furthermore, the use of conductive gel results in decent results (low impedances) on other subjects' custom-fit earpieces. This is somewhat surprising, since ear canals are believed to be more unique than fingerprints \underline{cite}. Finally, performance on FP1 was not as high as performance in the ear, despite FP1's popularity in past work on passthoughts \underline{cite}. This could be explained by the greater number of electrodes in the ear (compared to just one on FP1). Another explanation might be in the neural activity required to perform the tasks we chose. Future work might shed light on this issue.

@@

\section{Endnotes}

\begin{table}[h]
\centering
%% \begin{tabular}{llllll}
\begin{tabularx}{\textwidth}{llllll}
Task & Description & Stimuli? & Personal? & Eyes? & Imagery?\\
\hline
Breathe & Relaxed breathing with eyes closed & No & No & Closed & None\\
Breathe - Open & Relaxed breathing with eyes open & No & No & Open & None\\
Sport & Sport-related motor imagery & No & Yes & Closed & Motor\\
Song & Imagining hearing a song & No & Yes & Closed & Aural\\
Song - Open & Song task, with eyes open & No & Yes & Open & Aural\\
Speech & Imagining a spoken phrase & No & Yes & Closed & Aural\\
Listen \& Tone & Listening to a continuous tone & Yes & No & Closed & None\\
Listen - ASSR & Listening to noise modulated at 40 Hz & Yes & No & Closed & None\\
Face & Imagine a person's face & No & Yes & Closed & Visual\\
Sequence & Imagine face, a number, and word on timed cues & Yes & Yes & Open & Visual\\
\hline
\end{tabularx}
\caption{Properties of authentication tasks. We selected tasks with a variety of different properteries, but preferred tasks that did not require external stimuli, as the need to present such stimuli at authentication time could present challenges for usability, and user security.}
\label{tab:addlabel}%
\end{table}

>>>>

\subsection*{Abstract}
In this paper, we present a system that provides 3-factors of authentication (knowledge, posession and inherence) in a single step, using brain-based authentication via a custom-fit, in-ear EEG. Across all subjects, we achieve a mean 0\% false acceptance and 0.12\% false rejection rate with data from only one earpiece. In a preliminary test of an ``imposter'' spoofing attack, we find a 0\% false acceptance rate. Our results indicate that in-ear EEG could provide a discrete, convenient, multifactor authentication.

@@

In previous work, “one-step two-factor authentication” has been proposed as a
new approach to authentication that can provide the security benefits of two-
factor authentication without incurring the hassle costs of two-step verification.
By employing consumer-grade EEG (electroencephalogram) sensing
technologies, it was demonstrated in a 2013 passthoughts study that a user can
submit both a knowledge factor (i.e., secret thought) and an inherence factor
(i.e., brainwave signal unique to the individual) in a single step by performing a
single mental task \cite{Chuang2014}. Additionally, the robustness of this method against
impersonation attacks was demonstrated, including conditions where the attacker
may have learned the target’s secret thought and/or secret task \cite{Johnson2014}.

@@

\section{Related work}

\subsection{Passthought authentication}
The use of EEG as a biometric signal for user authentication has a short history.
In 2005, Thorpe et al. motivate and outline the design of a passthoughts system,
where, rather than typing a password, users authenticate 
by thinking of a passthought \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp \cite{Poulos2002,Marcel2007a,Palaniappan2008,Ashby2011}.
In 2013, one group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor \cite{Chuang2013b}. In particular, the
lack of signal diversity from multiple EEG channels can be overcome by allowing
the users to choose their own personalized passthoughts (e.g., sing their favorite
song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\textgreater \$10,000 USD)
and low usability (gel-based electrodes; aesthetic challenges of an EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task) \cite{Chuang2014}.

@@

\subsection{In-Ear EEG}
Research in in-ear EEG is only several years old. Nonetheless, the concept has
attracted a lot of attention because of the discreetness factor of in-ear EEG over
traditional scalp-based EEG. A research team at the Imperial College London
and Aarhus University published a landmark paper in 2011 that introduced the
concept of in-ear EEG, demonstrating for the first time the feasibility of recording
brainwave signals from within the ear canal
\cite{Looney2011}.
Follow-up work from the same
group demonstrated its ability to produce signal-to-noise ratios comparable to
those from conventional EEG electrode placements, robustness to common
sources of artifacts, and use in a brain-computer interface (BCI) system based on
auditory evoked potentials and visual evoked potentials
\cite{Looney2012a,Kidmose2013a,Kidmose2013b}.
United Sciences is currently developing a consumer ``hearable'' (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics \cite{UnitedSciences}.

\cite{curranpassthoughts} was the first to merge in-ear EEG with passthought authentication.

\subsection{Multi-Factor Authentication in a Single Step}
Behavioral authentication methods such as keystroke dynamics and speaker
authentication can be categorized as one-step two-factor authentication
schemes. In both cases, the knowledge factor (password or passphrase) and
inherence factor (typing rhythm or speaker’s voice) are employed \cite{Monrose1997}.
In contrast, the Nymi band supports one-step two-factor authentication via the inherence
factor (cardiac rhythm that is supposed to be unique to each individual) and the
possession factor (the wearing of the band on the wrist) \cite{Nymi}.
However, as far as we know, no one has proposed or demonstrated a one-step three-factor
authentication scheme, in which possession of a unique device also serves to authenticate the user.

\section{Methods}
\label{sec:org63d105c}
\subsection{{\bfseries\sffamily TODO} Manufacturing, materials}
\label{sec:orgb483219}
\subsection{Subjects}
\label{sec:orgaf30da9}
\subsection{{\bfseries\sffamily TODO} Tasks}

@@

\begin{table*}[t]
\centering
%% \begin{tabular}{llllll}
\begin{tabularx}{\textwidth}{llllll}
Task & Description & Stimuli? & Personal? & Eyes? & Imagery?\\
\hline
Breathe & Relaxed breathing with eyes closed & No & No & Closed & None\\
Breathe - Open & Relaxed breathing with eyes open & No & No & Open & None\\
Sport & Sport-related motor imagery & No & Yes & Closed & Motor\\
Song & Imagining hearing a song & No & Yes & Closed & Aural\\
Song - Open & Song task, with eyes open & No & Yes & Open & Aural\\
Speech & Imagining a spoken phrase & No & Yes & Closed & Aural\\
Listen \& Tone & Listening to a continuous tone & Yes & No & Closed & None\\
Listen - ASSR & Listening to noise modulated at 40 Hz & Yes & No & Closed & None\\
Face & Imagine a person's face & No & Yes & Closed & Visual\\
Sequence & Imagine face, a number, and word on timed cues & Yes & Yes & Open & Visual\\
\hline
\end{tabularx}
\caption{Properties of authentication tasks. We selected tasks with a variety of different properteries, but preferred tasks that did not require external stimuli, as the need to present such stimuli at authentication time could present challenges for usability, and user security.}
\label{tab:addlabel}%
\end{table*}

@@

We analyzed the EEG signals collected during the tasks using a support vector classifier (SVC). Since past work has shown that classification tasks in EEG-based BCI are linear \underline{Garrett2003a}, we used XGBoost, a tool for logistic linear classification \underline{Chen2016}.

@@

Counter to our expectations, we found that referencing on the same side as the electrodes improves classification accuracy, compared to referencing on the other side of the head, as is commonly done in EEG. \underline{Theory as to why?} Furthermore, the use of conductive gel results in decent results (low impedances) on other subjects' custom-fit earpieces. This is somewhat surprising, since ear canals are believed to be more unique than fingerprints \underline{cite}. Finally, performance on FP1 was not as high as performance in the ear, despite FP1's popularity in past work on passthoughts \cite{Chuang2013b}. This could be explained by the greater number of electrodes in the ear (compared to just one on FP1). Another explanation might be in the neural activity required to perform the tasks we chose. Future work might shed light on this issue.

@@

{\footnotesize \bibliographystyle{acm}
\bibliography{references}}
